### 2021-02-19 text/draft.tex
reconcilement

\begin{figure}
\centering
\includegraphics{images/chi_square_selection_fields.pdf}
\caption{The sample data set (\(n\) = 4'111) is not representative of
the raw Edoc data (\(n\) = 68'345) per department. Data field
\texttt{abstract}: \(\chi^2 (df=9) =\) 3'160.556, \(p < 0.001\); data
field \texttt{id\_number}: \(\chi^2 (df=9) =\) 4'209.0285,
\(p < 0.001\); data field \texttt{keywords}: \(\chi^2 (df=9) =\)
2'314.533, \(p < 0.001\). The data foundation is available at
\texttt{/edoc/analysis} and the chi-square statistic can be calculated
by calling \texttt{\_Analysis.chi\_square\_fit}.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{images/raw_sample_analysis.pdf}
\caption{The sample data set (\(n\) = 4'111) is significantly skewed
towards English (data field \texttt{language}: \(\chi^2 (df=3) =\)
1'441.414, \(p < 0.001\)) journal publications (data field
\texttt{type}: \(\chi^2 (df=10) =\) 52'519.743, \(p < 0.001\)) in the
sciences, medicine and economics (data field \texttt{department}:
\(\chi^2 (df=9) =\) 14'447.276\$, \(p < 0.001\)) from the 21st (data
field \texttt{date}: \(\chi^2 (df=7) =\) 35'878.493, \(p < 0.001\))
century as compared to the raw Edoc data (\(n\) = 68'345). The data
foundation is available at \texttt{/edoc/analysis} and the chi-square
statistic can be calculated by calling
\texttt{\_Analysis.chi\_square\_fit}.}
\end{figure}

@@

In what follows, I will describe how the cleaned keywords were
reconciled with Wikidata and YSO, and which additional steps for
refinement were undertaken. In total, 2'104 data transformation
operations were performed; the complete operation history is available
as \texttt{/keywords/operation\_history.json}.

A new project in OpenRefine was created and the data from
\texttt{keywords/keywords\_clean\_histogram.json} was imported. The
\texttt{keyword} column was then duplicated and reconciled with
Wikidata. Here the parameters were chosen as follows: reconcile against
no particular type, and auto-match candidates with high confidence.

@@

First, the suggestions to the top 500 keywords were manually verified.
These keywords account for 14'996 occurrences or 40.638\% of the total
occurrences and are constitute thus an effective lever.

@@

A second bias found was

After completing these steps to ameliorate the quality of the
reconciliation,

then suggests this scholarly article with high confidence.

However, certain types of concepts cannot qualify as subject terms

!! probabilistic random sampling to assess quality, weighted by
occurrences

!! calculate for each of annif yso mesh how many percent of keywords or
total occurrences are covered by reonciliated data: - Qid: 28789
occurrences, 9635 keywords - Yso: 14557 occurrences, 3185 keywords -
Mesh: 19775 occurrences, 4619 keywords

Before we do so, let us briefly look at the data in
\texttt{keywords/keywords\_clean.json}.

for a given input text, Annif assigns index terms from some controlled
vocabulary. Any gold standard employed to assess the quality of this
indexing must therefore make use of the same controlled vocabulary.

!! Mapping to controlled vocabularies; explain why this is necessary.

!! Does the OpenRefine project still exist? Export the list of changes!
http://127.0.0.1:3333/project?project=2501403769019\&ui=\%7B\%22facets\%22\%3A\%5B\%7B\%22c\%22\%3A\%7B\%22type\%22\%3A\%22list\%22\%2C\%22name\%22\%3A\%22qid\%22\%2C\%22columnName\%22\%3A\%22qid\%22\%2C\%22expression\%22\%3A\%22isBlank(value)\%22\%2C\%22omitBlank\%22\%3Afalse\%2C\%22omitError\%22\%3Afalse\%2C\%22selectBlank\%22\%3Afalse\%2C\%22selectError\%22\%3Afalse\%2C\%22invert\%22\%3Afalse\%7D\%2C\%22o\%22\%3A\%7B\%22sort\%22\%3A\%22name\%22\%7D\%2C\%22s\%22\%3A\%5B\%5D\%7D\%5D\%7D

>>>>

\begin{figure}
\centering
\includegraphics{images/chi_square_selection_fields.pdf}
\caption{The sample data set (\(n\) = 4'111) is not representative of
the raw Edoc data (\(N\) = 68'345) per department. Data field
\texttt{abstract}: \(\chi^2 (df=9) =\) 3'160.556, \(p < 0.001\); data
field \texttt{id\_number}: \(\chi^2 (df=9) =\) 4'209.0285,
\(p < 0.001\); data field \texttt{keywords}: \(\chi^2 (df=9) =\)
2'314.533, \(p < 0.001\). The data foundation is available at
\texttt{/edoc/analysis} and the chi-square statistic can be calculated
by calling \texttt{\_Analysis.chi\_square\_fit}.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{images/raw_sample_analysis.pdf}
\caption{The sample data set (\(n\) = 4'111) is significantly skewed
towards English (data field \texttt{language}: \(\chi^2 (df=3) =\)
1'441.414, \(p < 0.001\)) journal publications (data field
\texttt{type}: \(\chi^2 (df=10) =\) 52'519.743, \(p < 0.001\)) in the
sciences, medicine and economics (data field \texttt{department}:
\(\chi^2 (df=9) =\) 14'447.276\$, \(p < 0.001\)) from the 21st (data
field \texttt{date}: \(\chi^2 (df=7) =\) 35'878.493, \(p < 0.001\))
century as compared to the raw Edoc data (\(N\) = 68'345). The data
foundation is available at \texttt{/edoc/analysis} and the chi-square
statistic can be calculated by calling
\texttt{\_Analysis.chi\_square\_fit}.}
\end{figure}

@@

In this section, I will describe how the cleaned keywords were
reconciled with Wikidata and YSO, and which additional steps for
refinement were undertaken. In total, 2'104 data transformation
operations were performed; the complete operation history is available
as \texttt{/keywords/operation\_history.json}.

The data from \texttt{keywords/keywords\_clean\_histogram.json} was
imported into OpenRefine. The \texttt{keyword} column was then
duplicated and reconciled with Wikidata. Here the parameters were chosen
as follows: reconcile against no particular type, and auto-match
candidates with high confidence.

@@

First, the suggestions to the top 500 keywords were manually verified.
These keywords account for 14'996 occurrences or 40.638\% of the total
occurrences and thus constitute an effective lever.

@@

With these improvements in place, each keyword with a suggestion was
assigned a QID via the ``Add cloumns from reconciled values''-function
(and similar for YSO and MeSH identifiers). The data was then exported
and saved as \texttt{/keywords/keywords\_reference.json}. Keywords with
a QID now constitute 69\% of all keywords and 78\% of their total
occurrences in the sample data set, but these numbers are much lower for
YSO and MeSH identifiers (see Figure 4 for details).

\begin{figure}
\centering
\includegraphics{images/native_gold_standard.pdf}
\caption{The coverage of \texttt{/keywords/keywords\_reference.json} by
the controlled vocabularies of Wikidata (QID), Medical Subject Headings
(MeSH), and YSO (General Finnish Ontology). Both MeSH and YSO are
dependent on QID but independent of each other.}
\end{figure}

Let us now turn to the evaluation of the reconciliation: how many of the
suggestions were correct? This question was answered via random
sampling. The random sample (\(n=500\)) was created by calling
\texttt{\_Analysis.make\_random\_sample} on
\texttt{/keywords/reference\_keywords.json}; it is available at
\texttt{analysis/random\_keywords.json}. The sample was then imported
into OpenRefine and the judgement (1 for correct, 0 for incorrect) of
the manual verification was recorded in the column
\texttt{verification}. The data was then exported and is available at
\texttt{analysis/random\_keywords\_verified.csv}.

An analysis of this data shows that 53\% of the suggestions in the
random sample were correct with a 95\% confidence interval of 48.8\% to
57.3\%. We can therefore conclude that 8'507 \(\pm\) 692.1 of the 16'050
keywords in \texttt{/keywords/reference\_keywords.json} have correct
suggestions. Note that of the 235 incorrect suggestions in the random
sample, 188 were incorrect by default because they were missing a QID;
the share of non-empty yet incorrect suggestions is only 9.4\% in the
random sample meaning that the quality of the reconciliation is not as
disappointing as it might seem at first glance.

!! Talk about enrichment with yso and mesh
